---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r echo = TRUE}
unzip("activity.zip")
data <- read.csv("activity.csv")
data$date <- as.Date(data$date)
```


## What is mean total number of steps taken per day?

Here we find the sum of the number of steps for every day.  The summary is displayed below.
```{r echo = TRUE}
daily_sums <- with(data, tapply(steps, date, sum, na.rm = TRUE))
head(daily_sums)
```
Here we display the histogram of the number of steps taken every day.
```{r echo = TRUE}
hist(daily_sums)
```
```{r echo = TRUE}
mean_steps <- round(mean(daily_sums), 2)
median_steps <- median(daily_sums)
```
The mean of the total number of steps taken per day is `r mean_steps`. The median number of steps taken across all days is `r median_steps`.

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
