---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r echo = TRUE}
unzip("activity.zip")
data <- read.csv("activity.csv")
data$date <- as.Date(data$date)
```


## What is mean total number of steps taken per day?

Here we find the sum of the number of steps for every day.  The summary is displayed below.
```{r echo = TRUE}
daily_sums <- with(data, tapply(steps, date, sum, na.rm = TRUE))
head(daily_sums)
```
Here we display the histogram of the number of steps taken every day.
```{r echo = TRUE}
hist(daily_sums)
```
```{r echo = TRUE}
mean_steps <- round(mean(daily_sums), 2)
median_steps <- median(daily_sums)
```
The mean of the total number of steps taken per day is `r mean_steps`. The median number of steps taken across all days is `r median_steps`.

## What is the average daily activity pattern?
Below you see the graph of number of steps averaged across all days for every 5-minute interval time frame.
```{r echo = TRUE}
interval_mean <- with(data, tapply(steps, interval, mean, na.rm = TRUE))
plot(interval_mean, type='l', xlab="5-minute Interval Times", ylab="Average Number of Steps")
```
  
Below we find in which 5-minute interval has the average maximum number of steps across all days.
```{r echo = TRUE}
interval_mean[interval_mean == max(interval_mean)]
```
You can see interval 835 has an average of 206.2 steps over all days.


## Imputing missing values
To impute the missing steps values, here we load the "Hmisc" package and use the function "imputed" to calculate the mean of all the data to use for the NA values.

```{r echo = TRUE}
library(Hmisc)
new_data <- read.csv("activity.csv")
new_data$imputed_steps <- with(new_data, impute(steps, mean))
new_data$date <- as.Date(new_data$date)

```

Below is the new histogram with the imputed data filled in.
```{r echo = TRUE}
new_daily_sums <- with(new_data, tapply(imputed_steps, date, sum, na.rm = TRUE))
hist(new_daily_sums)
```

Below we find the new mean and median total number of steps. Then we take the difference from the originals.
```{r echo = TRUE}
new_mean_steps <- round(mean(new_daily_sums), 2)
mean_diff <- round(new_mean_steps - mean_steps, 2)
new_median_steps <- median(new_daily_sums)
median_diff <- new_median_steps - median_steps
```

The mean steps per day taken with the imputed values is `r new_mean_steps`. The difference from the original data is `r mean_diff`.  

The median steps per day taken with the imputed values is `r new_median_steps`. The difference from the original data is `r median_diff`.  

The impact of the imputed values is that both the mean and the median go up. You can also see from comparing the histograms that the middle of the day seems to be impacted more.

## Are there differences in activity patterns between weekdays and weekends?

First, create a function to tell weekend days from normal weekdays.  Then create a new varible that uses the function to tell the days.
```{r echo = TRUE}
day_type <- function(day) {
    if(day == "Saturday" | day == "Sunday") {
        return("weekend")
    } else {
        return("weekday")
    }
}
new_data$weekday <- weekdays(new_data$date)
new_data$day_type <- lapply(new_data$weekday, day_type)
new_data$day_type <- factor(new_data$day_type, levels=c("weekend", "weekday"))
summary(new_data)
```

Below we create a panel plot. This plot shows the average steps per day taken over every day for weekdays and for weekends separately.

```{r echo = TRUE}
new_data$interval <- as.factor(new_data$interval)
mean_by_day_type <- data.frame(with(new_data, tapply(steps, list(interval, day_type),
                                                     mean, na.rm = TRUE)))
mean_by_day_type$intervals <- levels(new_data$interval)
summary(mean_by_day_type)
par(mfrow = c(2, 1), mar = c(4, 4, 2, 0))
with(mean_by_day_type, plot(intervals, weekend, type='l', ylim=c(0, 250)))
with(mean_by_day_type, plot(intervals, weekday, type='l', ylim=c(0, 250)))
```

You can see by inspecting the y-axis bounds that earlier in the day, more steps are taken during the weekdays than the weekends.  And later in the day, more steps are taken during the weekends than the weekdays.

